---
title: "Masters"
output: html_document
---

Prepare workspace

```{r setup, include=FALSE}
getwd()
library(parallel)
library(doParallel)
library(caret)
library(MLeval)
library(ggplot2)
library(janitor)
library(mlr)
library(akmedoids)
```

Read data sets

```{r}
ace_vaxinpad <-
  read.csv("../Data/ace_vaxinpad/ace_vaxinpad_binary_centered.csv")
ace_vaxinpad_classes <-
  read.csv("../Data/ace_vaxinpad/ace_vaxinpad_classes.txt", header = F)
ace_vaxinpad_classes <- factor(ace_vaxinpad_classes[, 1], labels = c("No", "Yes"))

hiv_protease <-
  read.csv("../Data/hiv_protease/hiv_protease_binary_centered.csv")
hiv_protease_classes <-
  read.csv("../Data/hiv_protease/hiv_protease_classes.txt", header = F)
hiv_protease_classes <- factor(hiv_protease_classes[, 1], labels = c("No", "Yes"))
```

Split train / test

```{r}
set.seed(12)
trainIndACE <- createDataPartition(y = ace_vaxinpad_classes, p = 0.8, list = F)[, 1]
trainIndHIV <- createDataPartition(y = hiv_protease_classes, p = 0.8, list = F)[, 1]

ace_vaxinpad_train <- ace_vaxinpad[trainIndACE, ]
ace_vaxinpad_test <- ace_vaxinpad[-trainIndACE, ]

ace_vaxinpad_classes_train <- ace_vaxinpad_classes[trainIndACE]
ace_vaxinpad_classes_test <- ace_vaxinpad_classes[-trainIndACE]

hiv_protease_train <- hiv_protease[trainIndHIV, ]
hiv_protease_test <- hiv_protease[-trainIndHIV, ]

hiv_protease_classes_train <- hiv_protease_classes[trainIndHIV]
hiv_protease_classes_test <- hiv_protease_classes[-trainIndHIV]
```

Preprocessing function

```{r}
preprocessData <- function(data) {
  d <- remove_constant(data)
  corr <- cor(d, method = "spearman")
  featuresToRemove <- findCorrelation(corr, cutoff = 0.8)
  if (length(featuresToRemove) != 0) {
    d <- subset(d, select = -featuresToRemove)
  }
  return(d)
}
```

Preprocess data sets

```{r}
ace_vaxinpad_train <- preprocessData(ace_vaxinpad_train)
hiv_protease_train <- preprocessData(hiv_protease_train)

ace_vaxinpad_test <- subset(ace_vaxinpad_test, select = colnames(ace_vaxinpad_train))
hiv_protease_test <- subset(hiv_protease_test, select = colnames(hiv_protease_train))
```

Parallel training function

```{r}
createTrainControl <- function(metric, classes) {
  control <- trainControl(
    method = "cv",
    number = 10,
    index = createFolds(
      y = classes,
      k = 10,
      returnTrain = T,
      list = T
    ),
    classProbs = T,
    savePredictions = T,
    allowParallel = T
  )
  if (metric == "F") {
    control$summaryFunction <- prSummary
  }
  if (metric == "ROC")
    control$summaryFunction <- twoClassSummary
  
  return(control)
}

parallelTraining <- function(data, classes, model, metric) {
  no_cores <- detectCores() - 2
  cl <- makePSOCKcluster(no_cores)
  registerDoParallel(cl)
  
  output <- tryCatch({
    model <- caret::train(
      y ~ .,
      data = data.frame(data, y = classes),
      method = model,
      trControl = createTrainControl(metric = metric, classes = classes),
      metric = metric
    )
    return(model)
  }, warning = function(w) {
    print(w)
    return(NULL)
  }, error = function(e) {
    print(e)
    return(NULL)
  }, finally = {
    stopCluster(cl)
    registerDoSEQ()
  })
  return(output)
}
```

PCA based feature selection

```{r}
pcaFeatureSelection <- function(train, test, plot_label, method) {
  p <- prcomp(train)
  summ.p <- summary(p)
  var.explained <- summ.p$importance[2,]
  
  if (method == "total_var") {
    no_of_features <- which(cumsum(var.explained) > 0.8)[1]
    print(paste(
      no_of_features,
      "components account for 80% of total variance"
    ))
  }
  
  if (method == "elbow") {
    no_of_features <-
      floor(elbowPoint(x = c(1:length(p$sdev)), y = (p$sdev ^ 2))$x)
    print(paste(no_of_features, "components selected based on elbow method"))
  }
  
  gg <-
    ggplot(data = data.frame(x = c(1:length(p$sdev)), y = (p$sdev ^ 2)),
           mapping = aes(x = x, y = y)) +
    geom_point() +
    geom_line() +
    geom_vline(xintercept = no_of_features, color = "red")
  
  plot(
    gg + labs(
      title = "Screeplot",
      subtitle = plot_label,
      x = "Components",
      y = "Eigenvalues"
    )
  )
  
  train_return <- data.frame(p$x[, 1:no_of_features])
  
  test_return <-
    data.frame(predict(p, newdata = test)[, 1:no_of_features])
  
  return(list("train" = train_return, "test" = test_return))
}
```

Apply PCA based feature selection

```{r}
ace_vaxinpad_pca_var_list <-
  pcaFeatureSelection(
    train = ace_vaxinpad_train,
    test = ace_vaxinpad_test,
    plot_label = "ace vaxinpad",
    method = "total_var"
  )

ace_vaxinpad_pca_elbow_list <-
  pcaFeatureSelection(
    train = ace_vaxinpad_train,
    test = ace_vaxinpad_test,
    plot_label = "ace vaxinpad",
    method = "elbow"
  )

ace_vaxinpad_pca_var_train <- ace_vaxinpad_pca_var_list$train
ace_vaxinpad_pca_var_test <- ace_vaxinpad_pca_var_list$test

ace_vaxinpad_pca_elbow_train <- ace_vaxinpad_pca_elbow_list$train
ace_vaxinpad_pca_elbow_test <- ace_vaxinpad_pca_elbow_list$test
```

Filter based feature selection using gini impurity

```{r}
giniImportanceFeatureSelection <-
  function(data, classes, plot_label) {
    task <-
      makeClassifTask(data = data.frame(data, y = classes), target = "y")
    
    filterValues <-
      generateFilterValuesData(task, method = "ranger_impurity")
    
    sortedValues <- sort(filterValues$data$value, decreasing = T)
    
    elbow <-
      floor(elbowPoint(x = c(1:ncol(data)), y = sortedValues)$x)
    
    print(paste(elbow, "features selected based on elbow method"))
    
    gg <-
      ggplot(data = data.frame(x = c(1:ncol(data)), y = sortedValues),
             mapping = aes(x = x, y = y)) +
      geom_point() +
      geom_line() +
      geom_vline(xintercept = elbow, color = "red")
    
    plot(
      gg + labs(
        title = "Screeplot of gini importance",
        plot_label,
        x = "Features",
        y = "Gini importance"
      )
    )
    
    featuresSortedByImportance <-
      filterValues$data$name[order(filterValues$data$value, decreasing = T)]
    
    return(featuresSortedByImportance[1:elbow])
  }
```

Apply filter based feature selection

```{r}
ace_vaxinpad_gini_features <-
  giniImportanceFeatureSelection(ace_vaxinpad_train,
                                 ace_vaxinpad_classes_train,
                                 "ace vaxinpad")
ace_vaxinpad_train_gini <-
  subset(ace_vaxinpad_train, select = ace_vaxinpad_gini_features)
ace_vaxinpad_test_gini <-
  subset(ace_vaxinpad_test, select = ace_vaxinpad_gini_features)
```

Evaluation

ACE VAXINPAD

Original data set

```{r}
rf_ace_vaxinpad_original <-
  parallelTraining(
    data = ace_vaxinpad_train,
    classes = ace_vaxinpad_classes_train,
    model = "ranger",
    metric = "ROC"
  )

rf_ace_vaxinpad_original_eval <- evalm(data.frame(
  predict(rf_ace_vaxinpad_original,
          ace_vaxinpad_test,
          type = "prob"),
  ace_vaxinpad_classes_test
),
plots = c("r", "pr"))
```

PCA-reduced data set (using % of total variance method)

```{r}
rf_ace_vaxinpad_pca_var <-
  parallelTraining(
    data = ace_vaxinpad_pca_var_train,
    classes = ace_vaxinpad_classes_train,
    model = "ranger",
    metric = "ROC"
  )

rf_ace_vaxinpad_pca_var_eval <- evalm(data.frame(
  predict(rf_ace_vaxinpad_pca_var,
          ace_vaxinpad_pca_var_test,
          type = "prob"),
  ace_vaxinpad_classes_test
),
plots = c("r", "pr"))
```

PCA-reduced data set (using elbow method)

```{r}
rf_ace_vaxinpad_pca_elbow <-
  parallelTraining(
    data = ace_vaxinpad_pca_elbow_train,
    classes = ace_vaxinpad_classes_train,
    model = "ranger",
    metric = "ROC"
  )

rf_ace_vaxinpad_pca_elbow_eval <- evalm(data.frame(
  predict(
    rf_ace_vaxinpad_pca_elbow,
    ace_vaxinpad_pca_elbow_test,
    type = "prob"
  ),
  ace_vaxinpad_classes_test
),
plots = c("r", "pr"))
```

Gini importance-reduced data set

```{r}
rf_ace_vaxinpad_gini <-
  parallelTraining(
    data = ace_vaxinpad_train_gini,
    classes = ace_vaxinpad_classes_train,
    model = "ranger",
    metric = "ROC"
  )

rf_ace_vaxinpad_gini_eval <- evalm(data.frame(
  predict(rf_ace_vaxinpad_gini,
          ace_vaxinpad_test_gini,
          type = "prob"),
  ace_vaxinpad_classes_test
),
plots = c("r", "pr"))
```
